---
title: "Data Collection"
author: " By Claire M.K "
date: "September 19, 2017"
output: html_document
---


## The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv

and load the data into R. The code book, describing the variable names is here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FPUMSDataDict06.pdf



```{r setup, include=FALSE , warning=FALSE}
#Get data From URL
asUrl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06hid.csv"
download.file(asUrl, destfile ="asfile.csv" )
asdata <- read.csv("asfile.csv")

```


## How many properties are worth $1,000,000 or more?


```{r warning=FALSE}
#Subset the data, save int variable Qn1.
Qn1 <- subset(asdata, VAL == 24)
nrow(Qn1)
```



## Download the Excel spreadsheet on Natural Gas Aquisition Program here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx

Read rows 18-23 and columns 7-15 into R and assign the result to a variable called:
dat

```{r  echo=FALSE, Warning = FALSE}
# install package XLsx first, load the library 
library(xlsx)
library(xlsxjars)
#getin data from web
excurl = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2FDATA.gov_NGAP.xlsx"
#creating the file
urlf = file.path(getwd(),"DATA.gov_NGAP.xlsx")
download.file(excurl, urlf, mode = "wb") # download the file
dat <- read.xlsx(urlf, 1, rowIndex = 18:23, colIndex = 7:15 ) #read file into a dataframe

```


## What is the Value of

sum(dat$Zip*dat$Ext,na.rm=T)


```{r Warning = FALSE}
sum(dat$Zip * dat$Ext, na.rm = T)

```


## Read the XML data on Baltimore restaurants from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml


```{r warning=FALSE}
library(XML)
xmlurl <- "http://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Frestaurants.xml"
docu <- xmlTreeParse(xmlurl, useInternalNodes = TRUE)
RootNode <- xmlRoot(docu)
DocuName <- RootNode
```


## How many restaurants have zipcode 21231?

```{r warning=FALSE}
zipcode <- xpathSApply(RootNode, "//zipcode", xmlValue)
length(zipcode[zipcode==21231])
```


## The American Community Survey distributes downloadable data about United States communities. Download the 2006 microdata survey about housing for the state of Idaho using download.file() from here:

https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv


```{r}
pidurl <- "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv"
pid <- download.file(pidurl, destfile = "piddata.csv")
piddata <- read.csv("piddata.csv")
```

## Using the fread() command load the data into an R object DT
```{r warning=FALSE }
library(data.table)
DT <-fread(input = "https://d396qusza40orc.cloudfront.net/getdata%2Fdata%2Fss06pid.csv" ,
           sep = ",")

```


## Which is the fastest way to calculate the average value of the variable pwgtp15 broken down by sex using the data.table package?


```{r}
system.time(sapply(split(DT$pwgtp15,DT$SEX),mean))     

```

```{r}
system.time(mean(DT[DT$SEX==1,]$pwgtp15): mean(DT[DT$SEX==2,]$pwgtp15))
```


```{r}
system.time(mean(DT$pwgtp15,by=DT$SEX))
```


```{r}
system.time(DT[,mean(pwgtp15),by=SEX])
```
```{r}
system.time(tapply(DT$pwgtp15,DT$SEX,mean))
```

# END








