Red Wine Exploration by Claire M.K
========================================================
## Introduction:
> This report explores a dataset containing quality of red wine and 11 attributes of wine with a total of 1599 observations.The quality of the different wines was ranked using a scale of 0 to 10, with 10 being excellent and zero being worst. Each wine was ranked by atleast three wine experts

### Citation
The wine  dataset is public available for research. The details are described in [Cortez et al., 2009].

P. Cortez, A. Cerdeira, F. Almeida, T. Matos and J. Reis. Modeling wine preferences by data mining from physicochemical properties. In Decision Support Systems, Elsevier, 47(4):547-553. ISSN: 0167-9236.

Available at: [@Elsevier] http://dx.doi.org/10.1016/j.dss.2009.05.016 [Pre-press (pdf)] http://www3.dsi.uminho.pt/pcortez/winequality09.pdf [bib] http://www3.dsi.uminho.pt/pcortez/dss09.bib


```{r  echo=FALSE, message=FALSE, warning=FALSE, packages}
# Load all of the packages that you end up using
# in your analysis in this code chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk.
# This prevents the code from displaying in the knitted HTML output.
# You should set echo=FALSE for all code chunks in your file.
library(ggplot2)
library(gridExtra)
library(grid)
library('dplyr')
library('knitr') #we may need to remove this
library(ggthemes)

```
# Dataset

```{r echo=FALSE,Load_the_data}
# Load the Data
getwd()
setwd('C:/Users/Claire/Dropbox/Udacity/Project4 EDA')
Wine <- read.csv('wineQualityReds.csv')
#show the structure of the Wine dataframe
str(Wine)
#show summary statistics of the wine dataframe
summary(Wine)
```

# Univariate Plots
```{r echo=FALSE, message=FALSE, warning=FALSE}
p1=ggplot(aes( x=quality ), data =Wine)+ #histogram of quality
  geom_histogram(binwidth =0.5)+
  scale_x_continuous(breaks = seq(3,8,1.0))
#citric acid
p2=ggplot(aes( x=citric.acid ), data =Wine)+ #histogram of citric acid
  geom_histogram(binwidth = .01)+
  scale_x_continuous(breaks = seq(0.0,1.0,0.1))
#fixed.acidity
p3=ggplot(aes( x=fixed.acidity ), data =Wine)+ #histogram of fixed.acidity
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(4.50,16.00,1.0))
#residual.sugar
p4=ggplot(aes( x=residual.sugar ), data =Wine)+ #histogram of residual.sugar
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(1.0,15.50,2.0))
#density
p5=ggplot(aes( x=density ), data =Wine)+ #histogram of density
  geom_histogram(binwidth = 0.0001)+
  scale_x_continuous(breaks = seq(0.99,1.10,0.002))
#chlorides
p6=ggplot(aes( x=chlorides ), data =Wine)+ #histogram of chlorides
  geom_histogram(binwidth = 0.001)+
  scale_x_continuous(breaks = seq(.012,.611,.08))
#pH
p7=ggplot(aes( x=pH ), data =Wine)+ #histogram of pH
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(2.74,4.010,0.2))
#volatile.acidity
p8=ggplot(aes( x=volatile.acidity ), data =Wine)+ #histogram of volatile.acidity
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(.12,1.60,0.2))
#free.sulfur.dioxide
p9=ggplot(aes( x=free.sulfur.dioxide ), data =Wine)+ #histogram of free.sulfur.dioxide
  geom_histogram(binwidth = 1)+
  scale_x_continuous(breaks = seq(1,72,10))
#sulphates
p10=ggplot(aes( x=sulphates ), data =Wine)+ #histogram of sulphates
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(.33,2,.5))
#total.sulfur.dioxide
p11=ggplot(aes( x=total.sulfur.dioxide ), data =Wine)+ #histogram of total.sulfur.dioxide
  geom_histogram(binwidth = 1)
  scale_x_continuous(breaks = seq(4.50,16.00,1.0))
#alcohol
p12=ggplot(aes( x=alcohol ), data =Wine)+ #histogram of alcohol
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(8.4, 14.90,1))
grid.arrange(p1,p2,p3,p4,p5,p6,p7,p8,p9,p10,p11,p12,ncol=4)


```



From the above plots and statistics we can see that density, Quality, alcohol ,PH , volatile acidity, citric acid and fixed acidity appear to have a fairly normal distibution with the rest being right skewed. 
I will zoom in to see more of the individual plots and make a few transformations where necesary.

## A closer look at the acids

```{r echo=FALSE, message=FALSE, warning=FALSE, acid_plots}
#Univariate plots
a=ggplot(aes( x=citric.acid ), data =Wine)+ #histogram of citric acid
  geom_histogram(binwidth = .01)+
  scale_x_continuous(breaks = seq(0.0,1.0,0.1))
summary(Wine$citric.acid)#shows summary statistics
#fixed.acidity
summary(Wine$fixed.acidity)
a1=ggplot(aes( x=fixed.acidity ), data =Wine)+ #histogram of fixed.acidity
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(4.50,16.00,1.0 ))
#volatile Acidity
a2=ggplot(aes( x=volatile.acidity ), data =Wine)+ #histogram of volatile.acidity
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(.12,1.60,0.2))
a3=ggplot(aes( x=citric.acid ), data =Wine)+ #histogram of citric acid
  geom_freqpoly(binwidth = .01)+
  scale_x_continuous(breaks = seq(0.0,1.0,0.1))
a4=ggplot(aes( x=fixed.acidity ), data =Wine)+ #histogram of fixed.acidity
  geom_freqpoly(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(4.50,16.00,1.0 ))
#volatile Acidity
a5=ggplot(aes( x=volatile.acidity ), data =Wine)+ #histogram of volatile.acidity
  geom_freqpoly(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(.12,1.60,0.2))
grid.arrange(a,a1,a2,a3,a4,a5, ncol=3)

```

From the above plots we can see that citric acid seems to have a bimodal distribution , several peaks and a long tail, and what stands out most are the very many wines with no citric acid at all. since citric acid can add 'freshness' and flavor to wines.i will be taking a closer look at the wines with out citric acid later.

volatile acid seems normal but with a tail and several peaks with the highest peak at 0.6g / dm^3. Fixed acidity is also not very different from volatile, its seemingly normal shaped with a tail and several peaks.


```{r echo=FALSE, message= FALSE, warning= FALSE}
#density
summary(Wine$density)
ggplot(aes( x=density ), data =Wine)+ #histogram of density
  geom_histogram(binwidth = 0.0001)+
  scale_x_continuous(breaks = seq(0.99,1.10,0.002))
#quality
summary(Wine$quality)
ggplot(aes( x=quality ), data =Wine)+ #histogram of quality
  geom_histogram(binwidth =0.5)+
  scale_x_continuous(breaks = seq(3,8,1.0))
summary(Wine$pH)
ggplot(aes( x=pH ), data =Wine)+ #histogram of pH
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(2.74,4.010,0.2))
#alcohol
summary(Wine$alcohol)
ggplot(aes( x=alcohol ), data =Wine)+ #histogram of alcohol
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(8.4, 14.90,1))


```

From the above plots we see fairly normally distibuted plots with most wines having a density of 0.99 volume, Quality of 6, PH of 3.31 and alchol concetration of 10.20. alcohol seems to be leaning more to the right 

Its also important to note that about 82% of the wines are just average, with very few in the good rank the highest quality is 8 and lowest is 3.

### Residual Sugars
```{r echo=FALSE, message=FALSE, warning=FALSE, residual_sugar}
summary(Wine$residual.sugar)
r=ggplot(aes( x=residual.sugar ), data =Wine)+ #histogram of residual.sugar
  geom_histogram(binwidth = 1)+
  scale_x_continuous(breaks = seq(1.0,15.50,2.0))
r2=ggplot(aes( x=residual.sugar ), data =Wine)+ #histogram of residual.sugar
  geom_histogram(binwidth = .1)+
  scale_x_continuous(limits =c(0,7), breaks = seq(1,7,1))
grid.arrange(r,r2,ncol=2)
```

The residual.sugar plot was skewed and long tailed, I transformed the long tail data by changing the xlim to 7g / dm^3 from 15.50g / dm^3. The tranformed plot appears to be better representation of the residual.sugar distibution. Most wines have a concetration of sugar between 1.5g / dm^3 and 2.5g / dm^3

### Chloride

```{r echo=FALSE, message=FALSE, warning=FALSE, choride}
c=ggplot(aes( x=chlorides ), data =Wine)+ #histogram of chlorides
  geom_histogram(binwidth = 0.001)+
  scale_x_continuous(breaks = seq(.012,.611,.08))
c1= ggplot(aes(x=chlorides), data =Wine )+
  geom_freqpoly( binwidith =0.1)+ 
  scale_x_log10() #adding log10
c3=ggplot(aes( x=chlorides ), data =Wine)+ #histogram of chlorides
  geom_histogram(binwidth = 0.01)+
  scale_x_log10()
grid.arrange(c,c1,c3,ncol=2)
summary(Wine$chlorides)
```


The chloride plot was transformed to log10 for better visualisation and exploratioon.
most of the wines have a chloride concetration of 0.79g / dm^3

### Sulphurs

```{r echo=FALSE, message=FALSE, warning=FALSE}
#free.sulfur.dioxide
summary(Wine$free.sulfur.dioxide)
s=ggplot(aes( x=free.sulfur.dioxide ), data =Wine)+ #histogram of free.sulfur.dioxide
  geom_histogram(binwidth = 1)+
  scale_x_continuous(breaks = seq(1,72,10))
s1=ggplot(aes( x=log10(free.sulfur.dioxide) ), data =Wine)+ #histogram of free.sulfur.dioxide
  geom_histogram(binwidth = 0.05)+
  scale_x_continuous(breaks = seq(0,1.86,.5))
#sulphates
summary(Wine$sulphates)
s2=ggplot(aes( x=sulphates ), data =Wine)+ #histogram of sulphates
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(.33,2,.5))
s3=ggplot(aes( x=sulphates ), data =Wine)+ #histogram of sulphates
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(.33,1,.3))
#total.sulfur.dioxide
summary(Wine$total.sulfur.dioxide)
s4=ggplot(aes( x=total.sulfur.dioxide ), data =Wine)+#histogram of total.sulfur.dioxide
  geom_histogram(binwidth = 1)+
  scale_x_continuous(breaks = seq(6,289,70))

s5=ggplot(aes( x= log10(total.sulfur.dioxide+1) ), data =Wine)+ #histogram adding log
  geom_histogram(binwidth =0.02)+
  scale_x_continuous(breaks = seq(6,289,70))

grid.arrange(s,s2,s4,s1,s3,s5, ncol=3)


```

From the above plots we are able to see that all the suphurs have a mean that is higher than the median. i transformed the suphates plot by limiting the x axis to 1 for better visualization.
i also changed total and free sulphur dioxide to the log10 base. Total and free sulphur dioxide have similar distributions, i wonder if they are correlated.


```{r echo=FALSE, message=FALSE, warning=FALSE, alcohol}

```


### Wine Rating
```{r echo=FALSE }
# Transform the data type of quality to a factor
Wine$quality <- factor(Wine$quality, ordered = T)
# create a rating varible
Wine$rating <- ifelse(Wine$quality < 5, 'bad', ifelse(Wine$quality < 7, 'average', 'good'))
Wine$rating <- ordered(Wine$rating,levels = c('bad', 'average', 'good'))
summary(Wine$rating)


```
From the above statistics we can see that majority of the wines are just average only 13.5% of the wines are good.


```{r echo=FALSE,message=FALSE, warning=FALSE}

good <- subset(Wine, Wine$quality >6)
```

## Good Quality dataset

```{r echo=FALSE, message=FALSE, warning=FALSE,}

str(good)
summary(good)


```
# Good Quality -Univariate plots

```{r echo=FALSE , warning=FALSE , message= FALSE}

#citric acid
po2=ggplot(aes( x=citric.acid ), data =good)+ #histogram of citric acid
  geom_histogram(binwidth = .01)+
  scale_x_continuous(breaks = seq(0.0,.8,0.1))
#fixed.acidity
po3=ggplot(aes( x=fixed.acidity ), data =good)+ #histogram of fixed.acidity
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(4.50,16.00,1.0))
#residual.sugar
po4=ggplot(aes( x=residual.sugar ), data =good)+ #histogram of residual.sugar
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(1.0,9,2.0))
#density
po5=ggplot(aes( x=density ), data=good)+ #histogram of density
  geom_histogram(binwidth = 0.0001)+
  scale_x_continuous(breaks = seq(0.99,1.10,0.002))
#chlorides
po6=ggplot(aes( x=chlorides ), data =good)+ #histogram of chlorides
  geom_histogram(binwidth = 0.001)+
  scale_x_continuous(breaks = seq(.012,.36,.08))
#pH
po7=ggplot(aes( x=pH ), data =good)+ #histogram of pH
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(2.80, 3.78,0.2))
#volatile.acidity
po8=ggplot(aes( x=volatile.acidity ), data =good)+ #histogram of volatile.acidity
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(.12,0.92,0.2))
#free.sulfur.dioxide
po9=ggplot(aes( x=free.sulfur.dioxide ), data =good)+ #histogram of free.sulfur.dioxide
  geom_histogram(binwidth = 1)+
  scale_x_continuous(breaks = seq(3,54,10))
#sulphates
po10=ggplot(aes( x=sulphates ), data =good)+ #histogram of sulphates
  geom_histogram(binwidth = 0.01)+
  scale_x_continuous(breaks = seq(.39,1.36,.5))
#total.sulfur.dioxide
po11=ggplot(aes( x=total.sulfur.dioxide ), data =good)+ #histogram of total.sulfur.dioxide
  geom_histogram(binwidth = 1)
  scale_x_continuous(breaks = seq(7,289,70))
#alcohol
po12=ggplot(aes( x=alcohol ), data =good)+ #histogram of alcohol
  geom_histogram(binwidth = 0.1)+
  scale_x_continuous(breaks = seq(9.20, 14.90,1))
grid.arrange(po2,po3,po4,po5,po6,po7,po8,po9,po10,po11,po12,ncol=3)

```

looking at the plots with data from the wines with good quality, I still visualise outliers from residual sugar, free.sulphur.dioxide, sulphates, and total.sulphur.dioxide. the density, PH, chloride and alcohol plots seem to have a fairly normal distribution. Citric acid has a bimodal distribution. The histogram for free.Sulphur and total sulphur still  have a close resemblance. 



```{r echo=FALSE, message=FALSE, warning=FALSE}

```

```{r echo=FALSE, }


```



# Univariate Analysis

There are 1599 samples of red wine in the Wine dataset with 12 pysciochemical properties (fixed.acidity,volatile.acidity ,citric.acid, residual.sugar, chlorides , free.sulfur.dioxide ,total.sulfur.dioxide , density, pH ,sulphates, alcohol and quality). The variables are all of the numbers data type with the exeption of X and Quality who are intergers. 

The variable Quality is a measure of the quality of the sampled wines. using a scale of 0 to 10, with 10 being excellent and 0 being bad.


### Other observations:

* Most of the wines are of average quality ranked at 6 or 5 the min rank is 3 and max is 8.
* Most wines have an alcohol concetration between 1% and 11%.
* The density of most wines is 0.99 mg / dm^3
* Most wines have a PH of 3.33, but the PH of most good wines is at 3.27
* Residual sugar ,Free.sulphur.dioxide and total.sulphur.dioxide have very wide distributions
* 8.25% of the wines to not contain citric acid.
* Most wines have a concetration of Sulphates that is 0.62 g / dm3 but the good wines have a midian and mean of 0.74 g / dm3

### What is the structure of your dataset?
The main feature in the data set is the quality of the Wines. I’d like to determine which of the pysiochemical properties contribute or determine the quality of the wine. I also want to find out what quantities of the properties are needed to produce good quality wine.

### What is/are the main feature(s) of interest in your dataset?
The volume of alcohol,the PH of the wine and the quantity of residual sugar and sulphates seem to contribute to the quality of wines. More emphasis will be directed towards them

### What other features in the dataset do you think will help support your investigation into your feature(s) of interest?

Citric acid, volatile acidity and free sulphur dioxide will support my investigation. 

* PH is a measure of acidity so it makes sense to look at the acids
* Sulphates can contribute to free sulphur dioxide. 
* Its also important to note that  Citric acid contributes to the flavor and High volatile acidity quantities in wine can cause an unpleasant taste. 

### Did you create any new variables from existing variables in the dataset?
I changed the data type of quality to an ordered factor, and then created the variable rating with three distinct values.

1. Bad for the wines with a Quality value below 5
2. Average for the wines with a quality value of 5 and 6
3. Good for the wines with a quality value of 7 and 8

### Of the features you investigated, were there any unusual distributions? Did you perform any operations on the data to tidy, adjust, or change the form of the data? If so, why did you do this?

I log-transformed the right skewed features which were Chlorides, Free.Sulphur.dioxide, total.sulphur.dioxide and residual to help create a more normalized distribtution.

I also used the Scale_X_ to change the limits of residual and sulphates which were long tailed. changing the limits presented a better view of the plots 


# Bivariate Plots Section
```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. How did the feature(s) of interest vary with other features in the dataset?

### Did you observe any interesting relationships between the other features (not the main feature(s) of interest)?

### What was the strongest relationship you found?




# Multivariate Plots Section

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the investigation. Were there features that strengthened each other in terms of looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.

------

# Final Plots and Summary

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection
