# R Programming
## Programming Assignment 3

### Reading data into R

```{r warning=FALSE}
# read in the data
outcome = read.csv("outcome-of-care-measures.csv", colClasses = "character")
head(outcome, 2) #shows the first 2 rows and column headings
str(outcome) #shows structure of the dataframe
nrow(outcome) #returns number of observations
ncol(outcome)# number of variables
```


# 1. Plot the 30-day mortality rates for heart attack

Because the data was entered in as character we have to change the data type to numeric bbefore we can create the plots

```{r warning=FALSE}
# change datatype to numeric
outcome[,11] <- as.numeric(outcome[,11])
hist(outcome[,11])

```


# 2. Finding the best hospital in a state

Write a function called best that take two arguments: the 2-character abbreviated name of a state and an
outcome name. The function reads the outcome-of-care-measures.csv file and returns a character vector
with the name of the hospital that has the best (i.e. lowest) 30-day mortality for the specified outcome
in that state. The hospital name is the name provided in the Hospital.Name variable. The outcomes can
be one of "heart attack", "heart failure", or "pneumonia". Hospitals that do not have data on a particular
outcome should be excluded from the set of hospitals when deciding the rankings.
#### Handling ties. 

If there is a tie for the best hospital for a given outcome, then the hospital names should
be sorted in alphabetical order and the first hospital in that set should be chosen (i.e. if hospitals "b", "c",
and "f" are tied for best, then hospital "b" should be returned).

```{r warning=FALSE}
best <- function(state , outcome){
  # change data type from character to numeric
outcome[, 11] <- as.numeric(outcome[, 11]) # heart attack
outcome[, 17] <- as.numeric(outcome[, 17]) # heart failure
outcome[, 23] <- as.numeric(outcome[, 23]) # pneumonia

valid_outcomes <- c("heart attack", "heart failure", "pneumonia")
if (!state %in% outcome$State) {
  stop("invalid state")
  
} else if(!outcome %in% valid_outcomes) {
  stop("invalid outcome")
  
} else {
  if(outcome == "heart attack") {
    hosp_name <- helper(outcome, 11, state)
  } else if(outcome == "heart failure") {
    hosp_name <- helper(outcome, 17, state)
  } else {
    hosp_name <- helper(outcome, 23, state)
  }
  result <- hosp_name
  return(result)
}
  
}

```


### Test the function.

```{r warning=FALSE}
 #tests
best("TX", "heart attack")
best("TX", "heart failure")
best("MD", "heart attack")
best("AK", "pneumonia")
best("SC", "heart attack")
```

## 3 Ranking hospitals by outcome in a state
Write a function called rankhospital that takes three arguments: the 2-character abbreviated name of a
state (state), an outcome (outcome), and the ranking of a hospital in that state for that outcome (num).
The function reads the outcome-of-care-measures.csv file and returns a character vector with the name
of the hospital that has the ranking specified by the num argument. For example, the call
rankhospital("MD", "heart failure", 5)

would return a character vector containing the name of the hospital with the 5th lowest 30-day death rate
for heart failure. The num argument can take values "best", "worst", or an integer indicating the ranking
(smaller numbers are better). If the number given by num is larger than the number of hospitals in that
state, then the function should return NA. Hospitals that do not have data on a particular outcome should
be excluded from the set of hospitals when deciding the rankings.

### Handling ties.
It may occur that multiple hospitals have the same 30-day mortality rate for a given cause
of death. In those cases ties should be broken by using the hospital name. For example, in Texas ("TX"),
the hospitals with lowest 30-day mortality rate for heart failure are shown here.




```{r warning=FALSE}
rankhospital <- function(state, outcome, num = "best") {
    
    if(is.character(num)){
        if (num %in% c('best','worst')==FALSE){
            stop('Invalid Rank')
        } }
    
    data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
    
    if (state %in% data$State ==FALSE) { stop("invalid state")}
    
    if (outcome %in% c("heart failure","heart attack","pneumonia")==FALSE){
        stop("invalid outcome")}
    
   
 
    
    if (state %in% data$State & outcome %in% c("heart failure","heart attack","pneumonia")){
     if(outcome=="heart attack"){
         a<-data[,c(2,7,11)]
     }
     
     if(outcome=="heart failure"){
         a<-data[,c(2,7,17)]
     }
     
     if(outcome=="pneumonia"){
         a<-data[,c(2,7,23)]
     }
     
     b<-split(a,a$State)
     c<-b[[state]]
     c<-c[,c(1,3)]
     colnames(c)<-cbind('Hospital.Name',outcome)
     c[,2]<-as.numeric(c[,2])
     
     
     if (num=='best'){
                              
     d<- c[which(c[,2]==min(c[,2],na.rm=TRUE)),]

     hospitals<-d$Hospital.Name
     hospital<-sort(hospital)[1]
     
     }
     
     if (num=='worst'){
         
         d<- c[which(c[,2]==max(c[,2],na.rm=TRUE)),]
         
         hospitals<-d$Hospital.Name
         hospital<-sort(hospitals)[1]
         
     }
     
    if(is.numeric(num) & num < length(c$Hospital.Name)){
        
        
        l<-c[!is.na(c[,2]),]
        ll<-unique(sort(l[,2]))
        
         m<-c()
        
        for ( i in 1:length(ll)){
            temp<-which(l[,2]==ll[i])
            m<-rbind(m,l[temp,])
        }
        
        k<-m[num,2]
        p1<-which(m[,2]<k)
        p<-which(m[,2]==k)
        
        num2<-num-length(p1)
        
        hospitals<-sort(m[p,1])
        
        hospital<-hospitals[num2] 
        
    }
     
    if (is.numeric(num)){
        if (num > length(c$Hospital.Name)){ 
            return(NA)}}

    }
    
    hospital   
}
  
 

```

```{r warning=FALSE}
rankhospital("NC", "heart attack", "worst") 
rankhospital("WA", "heart attack", 7)
rankhospital("TX", "pneumonia", 10)
rankhospital("NY", "heart attack", 7)

```




# 4 Ranking hospitals in all states
Write a function called rankall that takes two arguments: an outcome name (outcome) and a hospital ranking
(num). The function reads the outcome-of-care-measures.csv file and returns a 2-column data frame
containing the hospital in each state that has the ranking specified in num. For example the function call
rankall("heart attack", "best") would return a data frame containing the names of the hospitals that
are the best in their respective states for 30-day heart attack death rates. The function should return a value
for every state (some may be NA). The first column in the data frame is named hospital, which contains
the hospital name, and the second column is named state, which contains the 2-character abbreviation for
the state name. Hospitals that do not have data on a particular outcome should be excluded from the set of
hospitals when deciding the rankings.

### Handling ties. 
The rankall function should handle ties in the 30-day mortality rates in the same way
that the rankhospital function handles ties.
The function should use the following template.




```{r warning=FALSE}
rankall <- function(outcome,num='best') {
    
    if(is.character(num)){
        if (num %in% c('best','worst')==FALSE){
            stop('Invalid Rank')
        } }
    
    data <- read.csv("outcome-of-care-measures.csv", colClasses = "character")
    
    
    if (outcome %in% c("heart failure","heart attack","pneumonia")==FALSE){
        stop("invalid outcome")}
    
   
 
    
    if (outcome %in% c("heart failure","heart attack","pneumonia")){
     if(outcome=="heart attack"){
         a<-data[,c(2,7,11)]
     }
     
     if(outcome=="heart failure"){
         a<-data[,c(2,7,17)]
     }
     
     if(outcome=="pneumonia"){
         a<-data[,c(2,7,23)]
     }}
    
    
    a[,3]<-as.numeric(a[,3])
    
    
    colnames(a)<-cbind('hospital','state','value')
    
    states<-sort(unique(a$state))
    
    b<-split(a,a$state)
    
    ans_state<-c()
    ans_hospital<-c()
    
    
    
    
    for (letter in states){

   
    g<-b[[letter]]
    
     g<-g[,c(1,3)]

    if (num=='best'){
        
        d<- c[which(g[,2]==min(g[,2],na.rm=TRUE)),]
        
        hospitals<-d$hospital
        hospital<-sort(hospitals)[1]
        
    }
    
    if (num=='worst'){
        
        d<- g[which(g[,2]==max(g[,2],na.rm=TRUE)),]
        
        hospitals<-d$hospital
        hospital<-sort(hospitals)[1]
        
    }
    
    if(is.numeric(num) & num < length(g$hospital)){
        
        
        l<-g[!is.na(g[,2]),]
        ll<-unique(sort(l[,2]))
        
        m<-c()
        
        for ( i in 1:length(ll)){
            temp<-which(l[,2]==ll[i])
            m<-rbind(m,l[temp,])
        }
        
        k<-m[num,2]
        p1<-which(m[,2]<k)
        p<-which(m[,2]==k)
        
        num2<-num-length(p1)
        
        hospitals<-sort(m[p,1])
        
        hospital<-hospitals[num2]
    }
    
    if (is.numeric(num)){
        if (num > length(g$hospital)){ 
            hospital<-NA}}
    
    ans_state<-c(ans_state,letter)
    ans_hospital<-c(ans_hospital,hospital)
    
    }
    
    df = data.frame(ans_hospital, ans_state)       
    colnames(df)<-c('hospital','state')
    
    df
}
  

```





```{r warning=FALSE}
r <- rankall("heart attack", 4)
as.character(subset(r, state == "HI")$hospital)

r <- rankall("pneumonia", "worst")
as.character(subset(r, state == "NJ")$hospital)

r <- rankall("heart failure", 10)
as.character(subset(r, state == "NV")$hospital)


```








